---
title: "Data integration"
author: "James Ashmore"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Setup

Set chunk options:

```{r knitr}
knitr::opts_chunk$set(
  autodep    = TRUE,
  cache      = TRUE,
  cache.path = "cache/01-data-integration.Rmd/",
  cache.lazy = FALSE,
  dev        = "png",
  error      = FALSE,
  message    = FALSE,
  warning    = FALSE
)
```

Load required packages:

```{r}
library(scater)
library(scran)
library(batchelor)
library(patchwork)
library(EnsDb.Mmusculus.v79)
library(org.Mm.eg.db)
library(msigdbr)
```

Read experiment data:

```{r}
sce <- list(
  NEG = readRDS("data/YS_GFP_NEG_C.rds"),
  POS = readRDS("data/YS_GFP_POS_D.rds")
)
```

Read variance data:

```{r}
dec <- list(
  NEG = metadata(sce$NEG)$modelGeneVar,
  POS = metadata(sce$POS)$modelGeneVar
)
```

## Preparation

Make feature names consistent:

```{r}
for (i in names(sce)) { rownames(sce[[i]]) <- rowData(sce[[i]])$gene_id }

for (i in names(sce)) { rownames(dec[[i]]) <- rowData(sce[[i]])$gene_id }
```

Reduce data and models to common features:

```{r}
ids <- Reduce(intersect, lapply(sce, rownames))

sce <- lapply(sce, "[", i = ids, )

dec <- lapply(dec, "[", i = ids, )
```

Per-batch scaling normalization:

```{r}
sce <- do.call(multiBatchNorm, sce)

dec <- do.call(combineVar, dec)

rownames(dec) <- ids
```

Blacklist genes for feature selection:

```{r}
use <- rownames(sce$NEG)[rowData(sce$NEG)$blacklist == FALSE]
```

Clear incompatible feature data:

```{r}
for (i in names(sce)) { rowData(sce[[i]]) <- rowData(sce[[i]])[, c("gene_id", "gene_name")]}
```

Select features for batchelor integration:

```{r}
hvg <- getTopHVGs(dec[use, ], var.field = "bio", var.threshold = 0, fdr.field = "p.value", fdr.threshold = 0.05)
```

## Correction

Perform mutual nearest neighbors correction:

```{r}
set.seed(1156433063)

mnn <- correctExperiments(
  "NEG" = sce$NEG,
  "POS" = sce$POS,
  subset.row = hvg,
  correct.all = TRUE,
  PARAM = FastMnnParam(merge.order = list("NEG", "POS"))
)
```

## Diagnostics

Check proportion of lost variance:

```{r}
metadata(mnn)$merge.info$lost.var
```

Cluster on the corrected PCA to obtain a partitioning of the cells:

```{r}
snn <- buildSNNGraph(mnn, type = "jaccard", use.dimred = "corrected")

com <- igraph::cluster_louvain(snn)

mnn$cluster <- factor(com$membership)
```

Tabulate number of cells from multiple batches within each cluster:

```{r}
tab <- table(Cluster = mnn$cluster, Batch = mnn$batch)

knitr::kable(tab)
```

Visualize the corrected PCA using a TSNE plot: 

```{r}
set.seed(1166689068)

mnn <- runTSNE(mnn, dimred = "corrected")

plotTSNE(mnn, colour_by = "batch", text_by = "cluster")
```

Visualize the corrected PCA using a UMAP plot: 

```{r}
set.seed(1027371528)

mnn <- runUMAP(mnn, dimred = "corrected")

plotUMAP(mnn, colour_by = "batch", text_by = "cluster")
```

## Processing

Make feature names unique:

```{r}
rownames(mnn) <- uniquifyFeatureNames(rowData(mnn)$gene_id, rowData(mnn)$gene_name)
```

Store combined variance model:

```{r}
rownames(dec) <- rownames(mnn)

metadata(mnn)$combineVar <- dec
```

Store highly variable genes:

```{r}
metadata(mnn)$getTopHVGs <- hvg
```

## Summary

### Output

Write experiment object to a file:

```{r}
saveRDS(mnn, file = "data/01-data-integration.rds")
```
